Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-11-20 papers/thesis-kirstie/thesis.tex
spelling
fixed fig reference
more eidts

SFMT is an extension of Fundamental Measure Theory (FMT)
that was developed for the Hard-sphere fluid, and which can be derived from
Scaled Particle Theory (SPT).  
%For the Hard Sphere fluid, there is another way to form an expression for 
%the excess Helmholtz free energy in terms of a homogeneous number density 
%and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
Scaled Particle Theory seeks a general expression for the excess chemical 
potential of a homogeneous Hard Sphere fluid which can then be integrated 
to find the excess Helmholtz free energy according to the relation
%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogenesous fluid corredsponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. 
\begin{align}
	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
\end{align}
Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemcial potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation, shown below, is known as Widom's Insertion 
Formula~\cite{Hansen}.
%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering
that at each point in the fluid, there is a probability that 
a hard sphere of radius R has space available to exist at that point
without overlapping with other spheres, or it does not, as shown in 
Figure~\ref{fig:p_overlap}.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{P_overlap.pdf}
    \caption{No energy is needed to place a sphere at a position where 
    there is no overlap such as shown by the black dot at the center of 
    the dotted region near the top. But it is not possible to place a 
    sphere at a position where an overlap occurs.}
    \label{fig:p_overlap}
  \end{figure}

Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, then becomes
\begin{align}
	\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
	         &= V_{diff} \\
             &= W_{rev}(R)
\end{align}
The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
that it takes no energy for a hard sphere to be placed at a point
in the fluid where it does not overlap with any other spheres 
($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
to place a sphere where it would overlap another sphere, as described 
in Figure~\ref{fig:p_overlap}.
In this case Equation~\ref{widoms-insertion-formula} becomes
%\begin{equation}\mu_{excess}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3) &  \mbox{for} & R_{\cavity}=R\end{equation}
\begin{align}
  \mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
\end{align}
where the radius of the cavity $R_{cavity}$ equals the radius of a hard sphere R,
and $nV_{Sphere}$ is the probability that a sphere cannot be placed at
some random point in the homogeneous fluid of number density, $n$, without
overlapping with another sphere.

%where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
%hard sphere, R, and $<e^{-\beta{V_{diff}}}>$ is evaluated using the 
%probablities that the cavity is either filled or not filled~\cite{Hansen}.
%For cavities that are much larger than one of the hard spheres, 
Forming a more general result by requiring 
continuity at the boundaries between equations written for fluids with 
cavities of a radius $\operatorname{R}_{cavity}$, that is larger, or very 
much larger than the hard sphere radius R, Scaled Partical Theory arrives at an 
expression~\cite{Hansen} for the excess Helmholtz 
free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
where
\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}

An expression for the excess Helmholtz free energy as a functional of the 
number density profile $n(\vec r)$ can be found using Fundamental Measure 
Theory (FMT)~\cite{rosenfeld1989}. FMT was developed for hard sphere fluids, and has been very 
sucessful at predicting the thermodynamic properties of inhomogeneous hard 
sphere fluids starting from an expression for the excess Helmholtz free 
energy obtained from Scaled Partical Theory~\cite{santos2012phi3}. 
Soft Fundamental Measure Theory (SFMT), which will be addressed later, 
extends FMT to fluids with soft potentials like that of the WCA fluid.

%\begin{equation}{\eta = \frac{NR_{3}}{V}}\end{equation}
%\begin{equation}{\eta = \left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}}\end{equation}
%\begin{equation}{\eta = n\frac{4}{3}\pi{R^3}=nR_{3}=n_{3}}\end{equation}
From Equation~\ref{Fexcess-SPT}, a quantity $\zeta_{3}$ can be defined as follows
\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=n\zeta_{3}}\end{equation}
where $\zeta_{3}$ is the volume of a sphere. Likewise,
\begin{equation}{\zeta_{2}=4\pi{R^2}}\end{equation}
\begin{equation}{\zeta_{1}=R}\end{equation}
\begin{equation}{\zeta_{0}=1}\end{equation}
where $\zeta_{2}$ is the surface area of a sphere, and $\zeta_{1}$ is the 
radius of a sphere. The quantities $\zeta_{3}, \zeta_{2}, \zeta_{1},$ 
and $\zeta_{0}$ are fundamental geometric measures of a sphere, and they 
will prove useful by defining a set of weighted densities $n_{i}=n\zeta_{i}$ 
where $n=\frac{N}{V}$. 
\begin{equation}\label{n3}{n_{3}=n\zeta_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
\begin{equation}\label{n2}{n_{2}=n\zeta_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
\begin{equation}\label{n1}{n_{1}=n\zeta_{1}=\left(\frac{N}{V}\right)R}\end{equation}
\begin{equation}\label{n0}{n_{0}=n\zeta_{0}=\left(\frac{N}{V}\right)}\end{equation}

\noindent The excess free energy in Eq~\ref{Fexcess-SPT} can be written 
in terms of the number density $\frac{N}{V}$
\begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
and using the expressions in Equations~\ref{n3} through~\ref{n0}, this becomes
\begin{equation}\label{FexfromSPT}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V}\end{equation}
which has the form 
\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
where 
\begin{align}
   \Phi_1 &= -n_{0}ln(1-n_{3}) \\
   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2} 
\end{align}

The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
densities given the relation $n_{i}=n\zeta_{i}$, where the weights are 
given by $\zeta_{0},\zeta_{1},\zeta_{2},\zeta_{3}$ respectively. These 
weights are valid for determining the excess free energy of a system of 
hard spheres with a homogeneous number density n. But these weights, and 
their related  weighted densities, can be written in a more general form 
allowing for their extention to a system with a spacially varying number 
density, that is, an inhomogeneous fluid.

Using a backwards Heavyside step function defined as
\begin{displaymath}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{displaymath}
the following expressions are obtained
\begin{equation}{\zeta_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
\begin{equation}{\zeta_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}

%\begin{equation}{\zeta_{1}=\frac{R_{2}}{4\pi{R}}=R}\end{equation}
\begin{equation}{\zeta_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
%\begin{equation}{\zeta_{0}=\frac{R_{2}}{4\pi{R}^2}=1}\end{equation}
\begin{equation}{\zeta_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}

The weighted densities can now be expressed as
\begin{equation}{\color{red} CHECK!~~\color{black}n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}}\end{equation} 
with weight functions
\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}

%\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
%\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{R}^2}}\end{equation}
\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
where the weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}

The weight functions can be related to the mayer function 
\begin{equation}{f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1}\end{equation} 
where r is the distance between the centers of two hard spheres. 
For a hard sphere fluid the interatomic potential $V(r)$ is given by 
\begin{displaymath}{V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.}\end{displaymath}
and so the mayer function for a hard sphere is given by 
\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
The mayer function can be written in terms of weight functions if two 
additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  
where the vector weights are defined as 
\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation}

Taking the derivative of the mayer function in Eq~\ref{f(r)step} will 
prove useful, and is given by
\begin{equation}\label{mayerder=w2convolution}{\frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}}\end{equation} 
where, for a hard sphere fluid the left side of Equation~\ref{mayerder=w2convolution} gives
\begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
\begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
Using the general relation for convolution, which is equal to the inverse 
Fourier Transform of a product of Fourier Transforms 
\begin{equation}{f \otimes g = \text{IFT[ FT}[f]\text{ x FT[g] }]}\end{equation} 
with
\begin{equation}{w_2(r)=\delta(r-R)}\end{equation} 
\begin{equation}{\text{Fourier Transform}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
the right side of Equation~\ref{mayerder=w2convolution} gives
\begin{equation}{w_2 \otimes w_2 = \text{Inverse Fourier Transform}[e^{-ikR}e^{-ikR}]}\end{equation} 
\begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
which matches Equation~\ref{mayerder}.

Incorporating into Equation~\ref{FexfromSPT} the two vector weights which 
were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
the excess free energy becomes
\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
where
\begin{align}
\Phi_1 &= -n_{0}ln(1-n_{3}) \label{eq:Phi1}\\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \label{eq:Phi2} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
\end{align}
and the tensor weight is given by
\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation}

The expression for the weighted densities can be written in a way that 
accomodates both homogenious fluids where the number denstiy n is constant, 
and inhomogeneous fluids where the number density becomes a function of 
the position $n(\vec{r})$.
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
%The equatons above were derived for a uniform fluid where the number density n is constant, 
%but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
For a spatially varying number density $n(\vec{r})$, the excess Free 
Energy becomes a functional of the number density profile $n(\vec{r})$ :
\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}

The equations developed so far have been for a hard sphere fluid. All of 
the weight function equations, as well as the equation for the deriviative 
of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
of a delta function characteristic of hard spheres. Soft Fundamental Measure 
Theory extends the Fundamental Measure Theory discussed so far to 
fluids with soft potenials by keeping the relationships between the weight 
functions, and which the weight functions satisfy the same, but replacing 
$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
A new set of weight functions can then be derived for a system of soft 
spheres which is what we do in forming our functional for the 
Weeks-Chandler-Anderson fluid as detailed in the chapter on Methods.

>>>>

SFMT is an extension of Fundamental Measure Theory (FMT)
that was developed for the Hard-sphere fluid, and which can be derived from
Scaled Particle Theory (SPT).  
%For the Hard Sphere fluid, there is another way to form an expression for 
%the excess Helmholtz free energy in terms of a homogeneous number density 
%and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
Scaled Particle Theory seeks a general expression for the excess chemical 
potential of a homogeneous Hard Sphere fluid which can then be integrated 
to find the excess Helmholtz free energy according to the relation
%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogenesous fluid corredsponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. 
\begin{align}
	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
\end{align}
Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemcial potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation, shown below, is known as Widom's Insertion 
Formula~\cite{Hansen}.
%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}

The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering
that at each point in the fluid, there is a probability that 
a hard sphere of radius R has space available to exist at that point
without overlapping with other spheres, or it does not, as shown in 
Figure~\ref{fig:p_overlap}. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{figs/P_overlap.pdf}
    \caption{No energy is needed to place a sphere at a position where 
    there is no overlap such as shown by the black dot at the center of 
    the dotted region near the top. But it is not possible to place a 
    sphere at a position where an overlap occurs.}
    \label{fig:p_overlap}
  \end{figure}
Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, then becomes
\begin{align} 
	\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
	         &= V_{diff} \\
             &= W_{rev}(R)  \label{workrev}
\end{align}
The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
that it takes no energy for a hard sphere to be placed at a point
in the fluid where it does not overlap with any other spheres 
($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
to place a sphere where it would overlap another sphere, as described 
in Figure~\ref{fig:p_overlap}.
In this case Equation~\ref{widoms-insertion-formula} becomes
%\begin{equation}\mu_{excess}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3) &  \mbox{for} & R_{\cavity}=R\end{equation}
\begin{align} \label{muexcess}
  \mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
\end{align}
where the radius of the cavity $R_{cavity}$ equals the radius of a hard sphere R,
and $nV_{Sphere}$ is the probability that a sphere cannot be placed at
some random point in the homogeneous fluid of number density, $n$, without
overlapping with another sphere. Combining  Equations~\ref{workrev} and \ref{muexcess} gives
\begin{align}
  W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
\end{align}
%where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
%hard sphere, R, and $<e^{-\beta{V_{diff}}}>$ is evaluated using the 
%probablities that the cavity is either filled or not filled~\cite{Hansen}.
%For cavities that are much larger than one of the hard spheres, 
Forming a more general result by requiring continuity of the functions, 
along with their first and second derivatives, for the reversible work
written for fluids with a cavitiy radius 
%$\operatorname{R}_{cavity}$ 
equal to, larger than, or very 
much larger than the hard sphere radius R, Scaled Partical Theory arrives 
at an expression~\cite{Hansen} for the excess Helmholtz free energy of 
a homogeneous Hard-Sphere fluid in terms of the packing fraction $\eta$ 
\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
where
\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}

%for the excess Helmholtz 
%free energy obtained from Scaled Partical Theory for a homogeneous 
%Hard-Sphere fluid, 
Fundamental Measure Theory~\cite{rosenfeld1989} (FMT) 
can be constructed from Equation~\ref{Fexcess-SPT}~\cite{santos2012phi3},  
and provides an expression for the excess Helmholtz free energy of an 
inhomogeneous 
Hard-Sphere fluid as a functional of the number density profile $n(\vec r)$. 
%FMT was developed for hard sphere fluids, and has been very sucessful 
%at predicting the thermodynamic properties of inhomogeneous hard sphere fluids. 
Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
is an extension of FMT and provides an expression for the excess Helmoholtz
free energy of fluids with soft potentials like that of the WCA fluid. 
It will be helpful to continue looking at FMT before addressing SFMT.

Fundamental Mearsure Theory begins by noticing that 
%\begin{equation}{\eta = \frac{NR_{3}}{V}}\end{equation}
%\begin{equation}{\eta = \left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}}\end{equation}
%\begin{equation}{\eta = n\frac{4}{3}\pi{R^3}=nR_{3}=n_{3}}\end{equation}
from Equation~\ref{Fexcess-SPT}, a quantity $G_{3}$ can be defined as follows
\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
where $G_{3}$ is the volume of a sphere. Likewise, by defining
quantities
\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=R}\end{equation}
\begin{equation}{G_{0}=1}\end{equation}
where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
%reproduced by introducing a set of weighted densities $n_{i}=nG_{i}$.
\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
geometric measures of a sphere, and are what give rise to the name
Fundamental Measure Theory. Introducing a set of weighted densities $n_{i}=nG_{i}$ 
given by
%and they will prove useful by introducing a set of weighted densities 
%$n_{i}=nG_{i}$ where $n=\frac{N}{V}$. 
\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
Equation~\ref{FexfromSPTinthermsofG} can be written as
%Writing Eq~\ref{Fexcess-SPT} in terms of the number density $n=\frac{N}{V}$
%\begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
%and using the expressions in Equations~\ref{n3} through~\ref{n0}, this becomes
\begin{align}\label{FexfromSPT}
    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
\end{align}
which has the form 
\begin{align}
    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
\end{align}
or for a spatially varying number density $n(\vec{r})$, 
\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
where 
\begin{align}
   \Phi_1 &= -n_{0}ln(1-n_{3}) \\
   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
\end{align}

The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
\begin{align}
    n_{i}=nG_{i}
\end{align}
These weights are valid for determining the excess free energy of a system of 
hard spheres with a homogeneous number density n. But these weights, and 
their related  weighted densities, can be written in a more general form 
allowing for their extention to a system with a spacially varying number 
density, that is, an inhomogeneous fluid. Using a backwards Heavyside step function defined as
\begin{displaymath}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{displaymath}
the following expressions are obtained
\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
%\begin{equation}{G_{1}=\frac{R_{2}}{4\pi{R}}=R}\end{equation}
\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
%\begin{equation}{G_{0}=\frac{R_{2}}{4\pi{R}^2}=1}\end{equation}
\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
where the weighted densities can now be expressed as
%CHECK
\begin{align}
    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
\end{align} 
or for a spatially varying number density $n(\vec{r})$,
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
with weight functions
\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
%\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
%\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{R}^2}}\end{equation}
\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
The weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
which will be important to keep in tact when moving to SFMT.

The weight functions can be related to the mayer function 
\begin{align}
     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
\end{align} 
where r is the distance between the centers of two hard spheres, and 
$V(r)$ is the interatomic potential between two hard spheres given by 
\begin{align}
    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
\end{align}
Evaluating Equation with $V(r)$ given by Equation,  
the mayer function for a Hard-sphere fluid is found to be
\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
which can be written in terms of the weight functions if two 
additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation} 
The mayor function can then be written in terms of the weight functions~\cite{Hansen}
\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}

Taking the derivative of the mayer function in Eq~\ref{f(r)step} provides 
another relation that will be important to keep in tact when moving to SFMT.
\begin{align}\label{mayerder=w2convolution}
     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
\end{align} 
This can easily be checked by evalating the left side for a Hard-sphere fluid as follows,
%where, for a hard sphere fluid the left side of Equation~\ref{mayerder=w2convolution} gives
\begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
\begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
and evaluating the right side by using the general relation for convolution which is equal to the Inverse 
Fourier Transform of a product of Fourier Transforms. 
\begin{equation}{f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]}\end{equation} 
Taking the Fourier Transform of the weight function $w_2(r)=\delta(r-R)$ 
%\begin{equation}{w_2(r)=\delta(r-R)}\end{equation} 
gives
\begin{equation}{\text{F.T.}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
and the right side of Equation~\ref{mayerder=w2convolution} becomes
\begin{equation}{w_2 \otimes w_2 = \text{I.F.T.}[e^{-ikR}e^{-ikR}]}\end{equation} 
\begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
which matches Equation~\ref{mayerder}.
